---
title: "Exploratory Analysis of Prince Lyrics"
author: "Ryan Jerving"
date: "April 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

This is a project based on the DataCamp tutorial by Debbie Liske, "Lyric Analysis with NLP & Machine Learning with R," @ <https://www.datacamp.com/community/tutorials/R-nlp-machine-learning>

The project aims to demonstrate three things:
* Text mining with R tidyverse tools
* Sentiment analysis, topic modeling, and natural language processing (NLP)
* Predictive analytics using machine learning tools

## Loading Libraries and Data Sets

First, I import any libraries needed to do these tasks, then pull in the `csv` that Liske created with Prince's lyrics, release year, and Billboard position for each song.

```{r load-libraries-and-data, message = FALSE, results= 'hide'}

# most of the libraries needed

library(dplyr) #data manipulation
library(ggplot2) #visualizations
library(gridExtra) #viewing multiple plots together
library(tidytext) #text mining
library(wordcloud2) #creative visualizations

# now the data, preventing R from converting strings to factors

prince_orig = read.csv("https://s3.amazonaws.com/assets.datacamp.com/blog_assets/prince_raw_data.csv", stringsAsFactors = FALSE)

```

## Intial Exploration

Let's see what the column headings are:

```{r explore-names, message = FALSE}

names(prince_orig)

```

From these, we'll only need row number `X`, the lyrics (`text`), `song` title, `album` title, release `year`, and `peak` Billboard, `US.Pop`, and `US.R.B` variables.

```{r selected-desired-columns}

prince <- prince_orig %>%
  select(lyrics = text, song, year, album, peak,
         us_pop = US.Pop, us_rnb = US.R.B)

# Check out the size of this table

dim(prince)

# That's 824 songs!

# And take a transposed glimpse of a sample track:

glimpse(prince[139,])

```